---
title: "Data analyses on annual findings of finances from 12 US states (Urban and Rural) during the stock market crash from 2008 to 2010"
output: html_notebook
---
# Install and load packages
  
```{r}
# Install using install.package function and library function

install.packages("tidyverse")
library(tidyverse)
```

# Used three data sets from CORGIS Dataset which have been transposed, combined, and transformed in excel for R purposes using read.csv()

```{r}
#read csv file

state_finances <- read.csv("../Project PROPEL/state_finances.csv")
```

* Note: only subset the following states: NY, CA, IL, TX, AZ, and PA & ME, VT, WV, MS, MT, and AR. These two groups will represent rural and urban states, which were chosen based on high number of population.

# Subset data from master dataset in CSV file

```{r}
# subset rows of states using filter function and subset columns using select function

states <- state_finances %>%
  filter(States %in% c("NEW YORK","CALIFORNIA", "ILLINOIS", "TEXAS", "ARIZONA", "PENNSYLVANIA", "MAINE", "VERMONT", "WEST VIRGINIA", "MISSISSIPPI", "MINNESOTA", "ARKANSAS")) %>%
  select(c("States", "Year", "Total.revenue", "Total.expenditure", "Taxes", "Debt.at.end.of.fiscal.year"))
```

# Create a side-by-side bar chart using ggplot

```{r}
# Create ggplot with the y-value of Total revenue

ggplot(states, aes(x=States, y=Total.revenue))+
  geom_bar(stat='identity', fill="forest green")+
  ylab("Total revenue") 

# Now, create another ggplot with the y-value of Total expenditure
ggplot(states, aes(x=States, y=Total.expenditure))+
  geom_bar(stat='identity', fill="forest green")+
  ylab("Total expenditure")

# Plot both two variables, Total.revenue and Total.expenditure

states_plot <- gather(states, key="measure", value="value", c("Total.revenue", "Total.expenditure"))

ggplot(states_plot, aes(x=States, y=value))+
  geom_bar(stat='identity', fill="forest green")+
  facet_wrap(~measure, ncol = 1) +
  labs(title = "Total Revenue and Expenditure across 12 US Urban and Rural States in 2008 stock market crash", x = "States", y = "Dollars in million")
```

# Plot a line graph

```{r}

```
